A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-15 tex/codatadp.tex
allow "first-order" parenthesis rules for \mu
minor

\paragraph{Step 1: Computing the Bidirectional Closure.}
Figure~\ref{fig:cc-rules} gives the basic rules of our calculus.
Together with the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, the rule \rn{Cong} computes the (upward) congruence closure,
while the rules \rn{Decomp} and \rn{Clash} together compute (downward) unification.
In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
and failures to unify equated terms are recognized by \rn{Clash}.
The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.

\begin{figure}[t]
\centering
\begin{tabular}{c}
\(
\inferrule{
  \tau \in \Data
  \quad
  t_1 : \tau
  \quad
  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
  \Val \ec{t_1} = (\vthinspace\mu x.\; t_2)
  \quad
  x \in FV( t_2 )
}{
  \bot
}
\)
\rn{Cyclic}
\\[5\jot]
\(
\inferrule{
 \tau \in \Codata
 \quad
 t_1, t_2 : \tau
 \quad
 %\tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
 \Val \ec{t_1} \vsim \Val \ec{t_2}
}{
 \Ec := \Ec,\: t_1 \teq t_2
}
\)
\rn{Bisimilar}
\end{tabular}
\caption{\,Rules for cyclicity and bisimilarity%.
}
\label{fig:ab-rules}
\end{figure}

\paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
In this step,
we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}, that is,
no nonredudant application of one of its rules can be applied to $\Ec$.
When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
Thus, we will in the following consider $\Ec$ as a set of equivalence classes of terms. 
For term $t \in \tEc$, we will write $\ec{t}$ to denote the equivalence class in $\Ec$ that containing $t$.

>>>>

\paragraph{Step 1: Computing the Bidirectional Closure.}
Figure~\ref{fig:cc-rules} gives the basic rules of the calculus.
Together with \rn{Refl}, \rn{Sym}, and \rn{Trans}, the \rn{Cong} rule computes the (upward) congruence closure,
whereas the \rn{Decomp} and \rn{Clash} rules together compute (downward) unification.
In the unification rules, additional equalities are inferred based on the injectivity of constructors by \rn{Decomp},
and failures to unify equated terms are recognized by \rn{Clash}.
The rule \rn{Conflict} recognizes when an equality and its negation both occur in $\Ec$, in which case $\Ec$ has no model.

\begin{figure}[t]
\centering
\begin{tabular}{c}
\(
\inferrule{
  \tau \in \Data
  \quad
  t_1 : \tau
  \quad
  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
  \Val \ec{t_1} = \mu x.\; t_2
  \quad
  x \in FV( t_2 )
}{
  \bot
}
\)
\rn{Cyclic}
\\[4\jot]
\(
\inferrule{
 \tau \in \Codata
 \quad
 t_1, t_2 : \tau
 \quad
 %\tpath{\Ec}{\ec{t_1}} = \tpath{\Ec}{\ec{t_2}}\sigma \neq \emptyset
 \Val \ec{t_1} \vsim \Val \ec{t_2}
}{
 \Ec := \Ec,\: t_1 \teq t_2
}
\)
\rn{Bisimilar}
\end{tabular}
\caption{\,Rules for cyclicity and bisimilarity%.
}
\label{fig:ab-rules}
\end{figure}

\paragraph{Step 2: Applying Cyclicity and Bisimilarity.}
In the second step,
we assume that $\Ec$ is saturated with respect to the rules in Figure~\ref{fig:cc-rules}---that is,
no rule can be applied to extend $\Ec$ further.
When this is the case, it is easy to see that $\Ec$ induces an equivalence relation over $\tEc$ such that two terms $t_1$ and $t_2$ are equivalent if and only if $t_1 \teq t_2 \in \Ec$.
Thus, we will in the following consider $\Ec$ as a set of equivalence classes of terms. 
For term $t \in \tEc$, we will write $\ec{t}$ to denote the equivalence class in $\Ec$ that contains $t$.
